Motivating Exploration in Reinforcement Learning

### 2020-04-16 paper/main.tex

\author{
  Bob Wei\\
  Department of Computer Science\\
  University of Waterloo\\
  \texttt{q25wei@uwaterloo.ca} \\
  % examples of more authors
  \And
  Akshay Patel \\
  Department of Computer Science\\
  University of Waterloo\\
  \texttt{akshay.patel@uwaterloo.ca} \\
  \AND
  Samir Alazzam \\
  Department of Computer Science\\
  University of Waterloo\\
  \texttt{q25wei@edu.uwaterloo.ca} \\
}

@@

Exploration is something that comes naturally to humans, almost as if there exists an innate sense of curiosity, craving to experience something new, something yet to be seen. Through observing human behaviour, it is clear that the ability to explore a domain is critical to the process of discovery and learning, regardless of the task being considered.

@@

We refer to \texttt{Breakout} and \texttt{Seaquest} as our dense reward environments. Specifically, \texttt{Seaquest} is most dense in that positive rewards are given for each \"kill\", which occurs very frequently due to the sheer number of enemies and the fact that the player's submarine is always firing projectiles at high velocity. In fact, positive rewards are already obtained in the early steps of training (as can be seen in Fig. \ref{fig:seaquest-breakout-reward}). \texttt{Breakout} provides a slightly sparser reward but is still denser than that of \texttt{Pong}; rewards are given for each brick broken which occurs every time the agent is able to return the ball. This means that in early training, when the agent is unable to consistently return the ball, there is no reward signal and so training performance is heavily dependent on an exploration factor.

>>>>

\author{
  Bob Wei\\
  Department of Computer Science\\
  University of Waterloo\\
  \texttt{q25wei@uwaterloo.ca} \\
  % examples of more authors
  \And
  Akshay Patel \\
  Department of Computer Science\\
  University of Waterloo\\
  \texttt{akshay.patel@uwaterloo.ca} \\
  \AND
  Samir Alazzam \\
  Department of Computer Science\\
  University of Waterloo\\
  \texttt{salazzam@uwaterloo.ca} \\
}

@@

Exploration is something that comes naturally to humans, almost as if there exists an innate sense of curiosity, craving to experience something new, something yet to be seen. Jean Piaget, a Swiss developmental psychologist, asserted that it is in human nature to be curious, inquisitive, and interested in novelty as they develop and acquire knowledge about their environment \cite{piaget}.Through observing human behaviour, it is clear that the ability to explore a domain is critical to the process of discovery and learning, regardless of the task being considered.

@@

We refer to \texttt{Breakout} and \texttt{Seaquest} as our dense reward environments. Specifically, \texttt{Seaquest} is most dense in that positive rewards are given for each ``kill'', which occurs very frequently due to the sheer number of enemies and the fact that the player's submarine is always firing projectiles at high velocity. In fact, positive rewards are already obtained in the early steps of training (as can be seen in Fig. \ref{fig:seaquest-breakout-reward}). \texttt{Breakout} provides a slightly sparser reward but is still denser than that of \texttt{Pong}; rewards are given for each brick broken which occurs every time the agent is able to return the ball. This means that in early training, when the agent is unable to consistently return the ball, there is no reward signal and so training performance is heavily dependent on an exploration factor.
